Practical Challenges with Web Service Evolution:\\A Review of Literature and Current Industry Practices

### 2013-05-30 paper.tex
cleaned up some typos, removed the soap/rest examples to save space

\subsection{SOAP Web Services} % (fold)
\label{sub:soap_web_services}
SOAP was designed in 1998 by Dave Winer, Don Box, Bob Atkinson, and Mohsen Al-Ghosein for Microsoft and is currently maintained by the World Wide Web Consortium. [http://en.wikipedia.org/wiki/SOAP] In brief, SOAP is an XML based protocol to communicate structured data and necessary headers between the consumers and publisher of a web service. Here is an example SOAP message representing a itinerary booking request:

\lstset{language=XML}
\begin{lstlisting}
<?xml version='1.0' ?>
<env:Envelope xmlns:env="http://www.w3.org/2003/05/soap-envelope">
 <env:Header>
  <m:reservation xmlns:m="http://travelcompany.example.org/reservation"
          env:role="http://www.w3.org/2003/05/soap-envelope/role/next"
           env:mustUnderstand="true">
   <m:reference>uuid:093a2da1-q345-739r-ba5d-pqff98fe8j7d</m:reference>
   <m:dateAndTime>2001-11-29T13:20:00.000-05:00</m:dateAndTime>
  </m:reservation>
  <n:passenger xmlns:n="http://mycompany.example.com/employees"
          env:role="http://www.w3.org/2003/05/soap-envelope/role/next"
           env:mustUnderstand="true">
   <n:name>Åke Jógvan Øyvind</n:name>
  </n:passenger>
 </env:Header>
 <env:Body>
  <p:itinerary
    xmlns:p="http://travelcompany.example.org/reservation/travel">
   <p:departure>
     <p:departing>New York</p:departing>
     <p:arriving>Los Angeles</p:arriving>
     <p:departureDate>2001-12-14</p:departureDate>
     <p:departureTime>late afternoon</p:departureTime>
     <p:seatPreference>aisle</p:seatPreference>
   </p:departure>
   <p:return>
     <p:departing>Los Angeles</p:departing>
     <p:arriving>New York</p:arriving>
     <p:departureDate>2001-12-20</p:departureDate>
     <p:departureTime>mid-morning</p:departureTime>
     <p:seatPreference/>
   </p:return>
  </p:itinerary>
  <q:lodging
   xmlns:q="http://travelcompany.example.org/reservation/hotels">
   <q:preference>none</q:preference>
  </q:lodging>
 </env:Body>
</env:Envelope>
\end{lstlisting}

Example SOAP Message of an airline reservation \[http://www.w3.org/TR/2007/REC-soap12-part0-20070427/\#L1161\]

@@

\subsection{RESTful Web Services} % (fold)
\label{sub:restful_web_services}
Being an XML based protocol, SOAP messages are often verbose. Also, the additional header information introduces communication overheads. On the other hand, RESTful web services provide a lightweight and flexible approach, where the HTTP protocol stack is leveraged without imposing constraints on the actual format of the messaging. The term Representational State Transfer (REST) was introduced by Roy Fielding \[https://en.wikipedia.org/wiki/Representational_state_transfer\#cite_note-Fielding-Ch5-1\] and is used interchangeably as RESTful web service or Web API. In brief, REST is an implementation for web services where user defined ``resources'' are created, accessed, modified and deleted utilizing the various features of the HTTP stack such as caching, security, layering and different request methods (GET, POST, PUT, DELETE etc.), content types, etc. For an example, the aforementioned SOAP message for itinerary booking can be translated to a JSON based RESTful service call as follows:

\lstset{language=json}
\begin{lstlisting}
{
  itinerary: {
    departure: {
      departing: 'New York'
      arriving: 'Los Angeles'
      departureDate: '2001-12-14'
      departureTime: 'late afternoon'
      seatPreference: 'aisle'
    },
    return: {
      departing: 'Los Angeles',
      arriving: 'New York',
      departureDate: '2001-12-20',
      departureTime: 'mid-morning'
    }
  },

  lodging: {
    preference: 'none'
  }

}
\end{lstlisting}

Example REST Message of Itenerary Booking

@@

@@

To identify a specific version of a resource, Leskey suggests using a unique identifier with an accompanying explanation of how to interpret it. The versioning identifiers should be used consistently and be readily accessible and understandable to it’s consumers. In addition to this, the author also recommends specifying a policy describing compatibility between the versions so the consumers clearly understand the impacts of version changes. Because a web service is a contract between the publisher and its consumers, the policies from both sides need to be considered for its evolution. As an example, the author mentioned a generic policy as follows:

@@

\subsubsection{Versioning:}
The author then identified the key reasons for changes between versions such as: a) service description changes, b) changes in functions, c) changes in the interaction mechanics, and d) change of known pre and post conditions. These changes would require new versions of the service as well as accompanying any explanation of the specific changes.

@@

\subsubsection{Implementation:}
Once an appropriate evolution and versioning strategy is selected, the required changes need to be implemented. In this section we have discussed the implementation approaches as suggested in the literature from the perspective of two levels, 1) protocol level and 2) source code level.

@@

In essence, the design technique can be described as follows: To start with, a web service can be developed as usual. Then, the interface of the web service is duplicated under a namespace V1. An implementation of this interface is created in the same V1 namespace that performs necessary translations on the data and delegates the methods to the original web service. This produces a snapshot of the V1 service which can be published for its consumers.
Now, to evolve to V2 of the service, the changes are made to the original web service. For any breaking change, it is compensated in a V1<-->V2 adapter. This V1 ←> V2 adapter is the first of the ``Chain of Adapters'' to be used for evolution. Once the changes are complete, the interface of the original web service is duplicated into a the V2 namespace. In the same namespace, an implementation of this V2 interface delegates the calls to the original web service, while the V1<-->V2 adapter targets this new implementation. This way, a new snapshot for the V2 of the service is created. Since the adapter compensates for all breaking changes from V1 to V2, existing V1 clients still continue to work. As the service evolves, introducing new adapters for each version would produce a chain of such adapters to ensure backward compatibility and code reuse while being able to serve multiple concurrent versions.

@@

% subsection systematic_literature_review (end)

@@

@@

\subsubsection{Documentation Challenges} % (fold)
\label{ssub:documentation_challenges}
A documentation of the web service is used by the consumer to understand and explore the features of the web service. Evolving web services introduce several challenges related to their documentation as follows:
\begin{itemize}
  \item How to produce version specific documentation and usage examples for a web service?
  \item How to provide an interactive explorer for different versions a web service?
  \item How to generate changelogs between versions?
\end{itemize}
% subsubsection documentation_challenges (end)

@@

\section{The References Section}\label{references}

In order to permit cross referencing within LNCS-Online, and eventually
between different publishers and their online databases, LNCS will,
from now on, be standardizing the format of the references. This new
feature will increase the visibility of publications and facilitate
academic research considerably. Please base your references on the
examples below. References that don't adhere to this style will be
reformatted by Springer. You should therefore check your references
thoroughly when you receive the final pdf of your paper.
The reference section must be complete. You may not omit references.
Instructions as to where to find a fuller version of the references are
not permissible.

We only accept references written using the latin alphabet. If the title
of the book you are referring to is in Russian or Chinese, then please write
(in Russian) or (in Chinese) at the end of the transcript or translation
of the title.

The following section shows a sample reference list with entries for
journal articles \cite{jour}, an LNCS chapter \cite{lncschap}, a book
\cite{book}, proceedings without editors \cite{proceeding1} and
\cite{proceeding2}, as well as a URL \cite{url}.
Please note that proceedings published in LNCS are not cited with their
full titles, but with their acronyms!

>>>>

\subsection{SOAP Web Services} % (fold)
\label{sub:soap_web_services}
SOAP was designed in 1998 by Dave Winer, Don Box, Bob Atkinson, and Mohsen Al-Ghosein for Microsoft and is currently maintained by the World Wide Web Consortium. [http://en.wikipedia.org/wiki/SOAP] In brief, SOAP is an XML based protocol to communicate structured data and necessary headers between the consumers and publisher of a web service.

@@

\subsection{RESTful Web Services} % (fold)
\label{sub:restful_web_services}
Being an XML based protocol, SOAP messages are often verbose. Also, the additional header information introduces communication overheads. On the other hand, RESTful web services provide a lightweight and flexible approach, where the HTTP protocol stack is leveraged without imposing constraints on the actual format of the messaging.

The term Representational State Transfer (REST) was introduced by Roy Fielding \[https://en.wikipedia.org/wiki/Representational_state_transfer\#cite_note-Fielding-Ch5-1\] and is used interchangeably as RESTful web service or Web API. In brief, REST is an implementation for web services where user defined ``resources'' are created, accessed, modified and deleted utilizing the various features of the HTTP stack such as caching, security, layering and different request methods (GET, POST, PUT, DELETE etc.), content types, etc.

@@

\section{Research Findings} % (fold)
\label{sec:research_findings}

\subsection{Literature Review} % (fold)
\label{sub:literature_review}

@@

\subsubsection{Versioning:}
To identify a specific version of a resource, Leskey suggests using a unique identifier with an accompanying explanation of how to interpret it. The versioning identifiers should be used consistently and be readily accessible and understandable to it’s consumers. In addition to this, the author also recommends specifying a policy describing compatibility between the versions so the consumers clearly understand the impacts of version changes. Because a web service is a contract between the publisher and its consumers, the policies from both sides need to be considered for its evolution. As an example, the author mentioned a generic policy as follows:

@@

The author then identified the key reasons for changes between versions such as: a) service description changes, b) changes in functions, c) changes in the interaction mechanics, and d) change of known pre and post conditions. These changes would require new versions of the service as well as accompanying any explanation of the specific changes.

@@

\subsubsection{Implementation:}
Once an appropriate evolution and versioning strategy is selected, the required changes need to be implemented. In this section we have discussed the implementation approaches as suggested in the literature from the perspective of two levels, 1) protocol level and 2) application level.

@@

In essence, the design technique can be described as follows: To start with, a web service can be developed as usual. Then, the interface of the web service is duplicated under a namespace V1. An implementation of this interface is created in the same V1 namespace that performs necessary translations on the data and delegates the methods to the original web service. This produces a snapshot of the V1 service which can be published for its consumers.

Now, to evolve to V2 of the service, the changes are made to the original web service. For any breaking change, it is compensated in a V1<-->V2 adapter. This V1 <--> V2 adapter is the first of the ``Chain of Adapters'' to be used for evolution. Once the changes are complete, the interface of the original web service is duplicated into a the V2 namespace. In the same namespace, an implementation of this V2 interface delegates the calls to the original web service, while the V1<-->V2 adapter targets this new implementation. This way, a new snapshot for the V2 of the service is created.

Since the adapter compensates for all breaking changes from V1 to V2, existing V1 clients still continue to work. As the service evolves, introducing new adapters for each version would produce a chain of such adapters to ensure backward compatibility and code reuse while being able to serve multiple concurrent versions.

@@

% subsection literature_review (end)

@@

% section research_findings (end)

@@

\subsection{Documentation Challenges} % (fold)
\label{ssub:documentation_challenges}
A documentation of the web service is used by the consumer to understand and explore the features of the web service. Evolving web services introduce several challenges related to their documentation as follows:
\begin{itemize}
  \item How to produce version specific documentation and usage examples for a web service?
  \item How to provide an interactive explorer for different versions a web service?
  \item How to generate changelogs between versions?
\end{itemize}
% subsubsection documentation_challenges (end)

@@
